---
title: "Analysis of the NYC Italian Menu Data"
author: "{Your name/group}"
---

## Introduction

The data consists of the 168 observations with 5 variables. Information on each variable is provided below:

- Price, price (in USD) of dinner (including one drink and a tip), 
- Food = customer rating of the food (out of 30)
- Décor = customer rating of the decor (out of 30)
- Service = customer rating of the service (out of 30)
- East = 1 (0) if the restaurant is east (west) of Fifth Avenue

We choose Price as our response and the remaining 4 variables: Food, Décor, Service, East to be used as predictors.

## Exploratory data analysis (Lecture 2)

We load in the New York City Italian menu data and output the first 6 observations.

```{r}
nyc_restaurant_data = read.csv("https://gattonweb.uky.edu/sheather/book//docs/datasets/nyc.csv", header = T)
head(nyc_restaurant_data)
```

{Comment on how we are going to use the variables. We going to use Food, Decor, Service as continuous? Does this make sense and why?}

We list summaries of the data. We see we have more restaurants in the dataset east of Fifth Avenue than West.
We don't seem to have any missing or problematic values.

```{r}
summary(nyc_restaurant_data)
```

{Comment on the summary}

We create boxplots of each predictor variable. We observe the food rating and service ratings have a larger spread. 

```{r}
par(mfrow = c(2, 2))
boxplot(nyc_restaurant_data$Decor, xlab = "Décor")
boxplot(nyc_restaurant_data$Food, xlab = "Food")
boxplot(nyc_restaurant_data$Service, xlab = "Service")
boxplot(nyc_restaurant_data$Price, xlab = "Price")
```

{Comment on the boxplots}

{Comment on unusual observations from the boxplots}

We plot a scatter plot matrix for each variable against each other to identify possible collinearity and other possible problematic relationships.

```{r}
nyc_restaurant_data_subset = nyc_restaurant_data[, c(3, 4, 5, 6)]
plot(nyc_restaurant_data_subset, col="cadetblue")
```

{Comment on the scatterplots of response versus predictors}

{Comment on the scatterplots of predictors versus predictors}

```{r}
cor(nyc_restaurant_data_subset) 
```

{Comment on the correlation}

Investigate the correlation in the predictors.

```{r}
fit_service = lm(Service ~ Food + Decor, 
         data = nyc_restaurant_data)
par(mfrow = c(2, 2))
plot(fit_service)
```

{Comment on diagnostics}

```{r}
summary(fit_service)
summary(fit_service)$r.squared
```

{Comment on summary output and coefficient of determination}

We investigate the relationship of only food and decor ratings.

```{r}
fit_food = lm(Food ~ Decor, 
         data = nyc_restaurant_data)
par(mfrow = c(2, 2))
plot(fit_food)
```

{Comment on diagnostics}

```{r}
summary(fit_food)
summary(fit_food)$r.squared
```

{ comment on coefficient of determination. Do we have evidence of high correlation? }


## Fitting the linear regression model (Lecture 3)

We fit the multiple linear regression model using foot, decor, service ratings and the categorical variable to identify if the restaurant is on the east or west of Fifth Avenue.
The summary of the linear regressin fit is provided below.

```{r}
fit = lm(Price ~ East + Food + Decor + Service, 
         data = nyc_restaurant_data)

par(mfrow = c(2, 2))
plot(fit)
```

{Comment on diagnostics}

```{r}
summary(fit)
summary(fit)$r.squared
```

{Comment on overall F test}

{Possibly comment on the coefficient of determination and t-tests}

Fit the reduced model and check diagnostics.

```{r}
fit_reduced = lm(Price ~ East + Food + Decor, data = nyc_restaurant_data)
par(mfrow = c(2, 2))
plot(fit_reduced)
```

{comment on diagnostics}

```{r}
summary(fit_reduced)
summary(fit_reduced)$r.squared
```

{ comment on the difference of the coefficient of determination of the reduced model and the full model }

{comment on if we should we reduce the model any further}

{Determine a final model}

```{r}
summary(fit_reduced)

confint(fit_reduced, level = .95)
```

{interpret coeffficients and comment on intervals}

We predict the new restaurant.

```{r}
predict(fit_reduced, data.frame(Food = 25, Decor = 25, East = 1), interval="confidence")

predict(fit_reduced, data.frame(Food = 25, Decor = 15, East = 1), interval="confidence")
```

{comment on the comparison}

{interpret the intervals}

## Conclusions

Our analysis provides evidence to maximumize the menu price, the location of the restaurant should 
{Comment on east or west of Fifth Avenue?}

{Comment on how much this location benefits the menu price}


{Comment on which variables should be focused on to maximize the menu price. Take into account correlation of the predictors in your interpration}


{ Do you recommend a price premium for “setting a new standard for high-quality service in Manhattan” for Italian restaurants? why or why not?}
